---
title: "A5"
author: "SUS Teaching Team"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading the required libraries. 
```{r}
library(censusapi)
library(dplyr)
library(ggplot2)
library(tidyr)
```

Enter your census key here. 
```{r}
Sys.setenv(CENSUS_KEY="ENTER_YOUR_KEY_HERE")
```

You can search tk for lines that have to be changed. To perform this for multiple counties, you can simply wrap the code below in a for statement. You will need a list of the counties CA county FIPS codes. Note that at the moment, the code only works for data tabulated similarly to income, B19001. That is, each name is a variable of interest by some categorie (e.g. race) and within each name are the outcome. This is income tiers within the income by race. Refer to the website [here](https://api.census.gov/data/2016/acs/acs5/groups.html). You can initialize an empty list, or vector, to hold the final skew for each county. Refer to line 35 for an example of initializing an empty vector of a certain length.  


```{r}
county <- "001"

# Enter the number of categories here, likely the number of letters that span the variable of interest. 
num_cats <- 7
# Initializing the list that will hold the data. 
groups <- vector("list", num_cats)


for (i in 1:num_cats) {
  
  # First group - tk
  var_name <- paste("group(B19001",LETTERS[i],")", sep = "")
  
  
  counties_race <- getCensus(name = "acs/acs5", vintage = 2016, vars = c("NAME",var_name), region = paste("county:",county, sep = ""), regionin = "state:06")
  
  group_name <- paste("B19001",LETTERS[i], sep = "")
  
  # Need to adjust based on number of outcomes of interest - tk
  # For example, there's 16 outcomes for income + 1 for the total. Note that the second paste statement can be deleted if there are fewer than 10 outcomes. 
  # , paste(group_name,"_0", 10:17, "E", sep = "")
  vars_of_int_loop <- c(paste(group_name,"_00", 1:9, "E", sep = ""), paste(group_name,"_0", 10:17, "E", sep = ""))
  
  var_name <- paste("estimate", LETTERS[i], sep = "_")
  
  groups[[i]] <- select(counties_race,c("state","county", "NAME", vars_of_int_loop)) %>% gather(variable, value, -c("state","county", "NAME")) %>% group_by(variable) %>% summarise("estimate" = sum(value)) %>% select(estimate)
  
}

by_race <- do.call(cbind, groups)

num_tiers <- nrow(by_race)

percents <- t(apply(by_race, 1, function(row_i) return(row_i/sum(row_i))))

# Next is each entry subtracted from the first row. 
# We're telling R which row we don't want. 
skew <- t(apply(percents[-1, ], 1, function(row_i) return(row_i - percents[1,])))

# Initializing the empty final skew matrix. 
final_skew <- matrix(NA, ncol = ncol(skew), nrow = nrow(skew))

# Each entry of next is skew entry times sum from corresponding bracket in by race divided by total in by race. 
for (j in 1:(num_tiers-1)){
  final_skew[j,] <- abs(skew[j,]*sum(sum(by_race[(j+1),])/sum(by_race[1,])))
}
sum(final_skew)*100/2
```




